\mytitle

### 2020-04-06 paper.tex
commit.

\makeatletter
\providecommand{\doi}[1]{%
	\begingroup
	\let\bibinfo\@secondoftwo
	\urlstyle{rm}%
	\href{http://dx.doi.org/#1}{%
		doi:\discretionary{}{}{}%
		\nolinkurl{#1}%
	}%
	\endgroup
}
\makeatother
%\usepackage{natbib}
% encoding and languages
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[ngerman,english]{babel}
\usepackage{wrapfig}
\usepackage{times}
\usepackage{paralist}
\usepackage{graphicx, color}
%\usepackage{subfigure}
\usepackage{subfig}
\usepackage{subfloat}
\usepackage{booktabs}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{paralist}
\usepackage{amssymb, amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsopn}
\newtheorem{myProb}{Problem}
\newtheorem{mythm}{Theorem}
\newtheorem{myproof}{Proof}
\newtheorem{mysketch}{Proof Sketch}
%\usepackage[ruled]{algorithm2e}
%\usepackage[colorinlistoftodos]{todonotes}
%\usepackage{qtree}
\usepackage{tikz, tikz-qtree}
\usepackage[ruled]{algorithm2e}
%\usepackage{Algorithmic}
\usepackage{times}
\usepackage{microtype}
\usepackage{url}
\usepackage{balance}
\let\proof\relax 
\let\endproof\relax
\usepackage{multirow}
\usepackage{relsize}
\usepackage{caption}
\usepackage{color, colortbl}
\usepackage{array, booktabs, ragged2e}
\usepackage{makecell}
\usepackage{tikz}
\usepackage{textpos}

@@

\begin{abstract}
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem concerns
selecting turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic constraint as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model excels at finding high quality solutions very quickly while the QCOP model is able to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

We revisit 
the wind farm layout optimization (WFLO) problem 
that was first considered 
in~\citet{MOSETTI1994105}. 
In WFLO, we aim to place wind turbines 
in a predefined area to maximize 
the total generated power. 
Turbine location decisions    
influence
the
total power production due to  
interference effects, aka wakes, generated by upstream 
turbines. Furthermore, 
there are proximity constraints between turbines
that prevent from turbines to be placed next to each other.
These inter-turbine effects 
are naturally captured using quadratic constraints and objective
functions. Several exact and heuristic optimization approaches have 
been proposed to solve
the WFLO problem~\cite{turner2014new,Zhang2014}. However, these
approaches represented the quadratic effects and constraints
using linear models that approximate the real dependencies.

@@

In this part, we first present the
physical model of a wind farm that we use throughout the paper. 
Subsequently, we first provide an overview of 
existing mathematical
optimization methods for WFLO, focusing
on their similarities and differences
compared to our approach.

\subsection{Wind Farm Model}

We consider a square shaped wind farm that we model 
as an $N\times N$ grid
with $N$ being the number of cells in each axis. 
Each cell has the area of $c^2$ (in meters), i.e.,
if $c=100$ meters and $N=20$, the farm area will be $2km\times 2km$.
In this work, we assume that the terrain is flat and that 
we have no constraints related to the terrain.
Furthermore, we assume a sterile environment and do not consider the influence of noise
on the wind farm's surroundings. The wind farm literature treats the more complex
cases of non-flat terrain and environmental considerations in~\cite{} and~\cite{}, respectively.
In our work, we introduce a novel optimization approach that can be extended,
in the future, to include 
such considerations.

In principle, a higher number of turbines placed in the wind farm produces more energy. However,
we must consider two types of interference between the turbines.

First, 
turbines cannot be placed at high proximity to each other. 
Specifically, the turbines must be placed at least five rotor diameters apart from each other. Depending on the cell size, we shall formulate the corresponding constraints as part of the optimization model. 
For example, assuming that a turbine is placed
at the center of the cell (cell size of $100$ meters) and for rotor diameter of $40$ meters, 
we must ensure that two turbines will not placed in adjacent cells.

Second, turbines influence each other through an effect called \emph{wake}. Wakes cause reduction of the effective power produced by a turbine due to upstream turbines changing airflow dynamics.

\todo{Explain the wake model - single wake and linear superposition. Explain that it is less accurate but leads to a quadratic objective. We then assess it using sum of squares. Explain that there are $m$ turbines to be placed. use Zheng and use Mosetti.}

\todo{Wind direction}

@@

The allocation of turbines in a wind farm was first considered 
as an optimization problem in~\cite{MOSETTI1994105}. Since
then, WFLO has caught the attention of optimization researchers. 
The WFLO problem can either be solved \emph{exactly} using
techniques from operations research (e.g., linear programming, 
mixed-integer programming, stochastic programming, and constraint programming)\cite{Zhang2014,turner2014new}
or solved \emph{approximately} using various meta-heuristics (e.g., genetic algorithms and local search) \cite{MOSETTI1994105}.
The main advantage of the former techniques
is that they provide guarantees on solution quality, 
while the latter methods generate 
quick and sufficiently accurate solutions. In this work, we compare 
exact and approximate methods for WFLO. Both use 
quadratic optimization models... \todo{add more variations and explain that we
	focus on one to demonstrate the usefulness of quadratic models}

@@

In this part, 
we formulate the WFLO problem
as two quadratic optimization problems: a
quadratic constrained optimization (QCOP) problem and 
a quadratic unconstrained binary optimization (QUBO) problem. 
To this end,
we follow standard WFLO definitions
and notation presented in~\cite{Zhang2014}. First, we present WFLO's inputs and decision variables.
Subsequently, we present the mathematical formulation of the two problems.

\paragraph{WFLO Inputs and Decision Variables} Let $x_i$ be a binary decision variable that represents whether a 
turbine is positioned at location $i \in \mathcal{N}$ with $\mathcal{N}$ being the set 
of $n$ possible locations and 
$m$ being the total number of required turbines. Note that we assume a discrete grid with 
a finite number of possible positions. We 
write $x$ as shorthand for $x = (x_1,\ldots, x_n)$. Furthermore,
we denote by $\mathcal{E} \subseteq \mathcal{N}\times \mathcal{N}$
the set of location pairs that cannot simultaneously host turbines 
due to proximity constraints. 
Let $u_{id}$ and $u_{id, \infty} $ 
be the wind speeds at turbine location $i$ for wind regime $d \in \mathcal{D}$
with and without interference from other turbines due to wake effects, respectively. A wind regime is a combination 
of wind speed and direction. 
The probability of wind regime 
$d$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. For example, 
regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind.

Next, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
the set of downstream 
turbine locations 
given that a turbine is placed at cell $i$ for wind regime $d$ (i.e., turbines placed in these locations 
will be influenced by a turbine placed at $i$). Finally,
we denote by $u_{ijd}$ the wind speed at turbine $j$ due to a single wake from upstream turbine $i$ with $j \in \mathcal{U}_{id}$. 
The following objective function  
represents the total expected energy of a wind farm given placement decisions 
$x$: \begin{equation}
f(x) = \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \label{ObjFunc}\\
%&s.t.:& \nonumber\\
%&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
%&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
%&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{equation}

@@

To 
maximize $f(x)$ and satisfy the number of turbines and proximity 
constraints we can 
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{eqnarray}

Solving QCOP with Gurobi. Explain.

@@

\paragraph{WFLO Instances}

@@

>>>>

\makeatletter
\providecommand{\doi}[1]{%
	\begingroup
	\let\bibinfo\@secondoftwo
	\urlstyle{rm}%
	\href{http://dx.doi.org/#1}{%
		doi:\discretionary{}{}{}%
		\nolinkurl{#1}%
	}%
	\endgroup
}
\makeatother
%\usepackage{natbib}
% encoding and languages
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[ngerman,english]{babel}
\usepackage{wrapfig}
\usepackage{times}
\usepackage{paralist}
\usepackage{graphicx, color}
%\usepackage{subfigure}
\usepackage{subfig}
%\usepackage{subfloat}
\usepackage{booktabs}
\usepackage{caption}
%\usepackage{subcaption}
\usepackage{paralist}
\usepackage{amssymb, amsfonts}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsopn}
\newtheorem{myProb}{Problem}
\newtheorem{mythm}{Theorem}

\newtheorem{myproof}{Proof}
\newtheorem{mysketch}{Proof Sketch}
%\usepackage[ruled]{algorithm2e}
%\usepackage[colorinlistoftodos]{todonotes}
%\usepackage{qtree}
\usepackage{tikz, tikz-qtree}
\usepackage[ruled]{algorithm2e}
%\usepackage{Algorithmic}
\usepackage{times}
\usepackage{microtype}
\usepackage{url}
\usepackage{balance}
\let\proof\relax 
\let\endproof\relax
\usepackage{multirow}
\usepackage{relsize}
\usepackage{caption}
\usepackage{color, colortbl}
\usepackage{array, booktabs, ragged2e}
\usepackage{makecell}
\usepackage{tikz}
\usepackage{textpos}

@@

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem concerns
selecting turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is natural to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model excels at finding high quality solutions very quickly while the QCOP model is able to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

We revisit 
the wind farm layout optimization (WFLO) problem 
that was first considered 
in~\citet{MOSETTI1994105}. 
In WFLO, we aim to place wind turbines 
in a predefined area to maximize 
the total generated power. 
Turbine location decisions    
influence
the
total power production due to  
interference effects, aka wakes, generated by upstream 
turbines. Furthermore, 
there are proximity constraints between turbines
that prevent from turbines to be placed next to each other.
These inter-turbine effects 
are naturally captured using quadratic constraints and objective
functions. Several exact optimization approaches have 
been proposed to solve
the WFLO problem~\cite{turner2014new,Zhang2014}. However, these
approaches represented the quadratic effects and constraints
using linear models that approximate the true dependencies.

@@

In this part, we start
by presenting the
physical model 
of a wind farm that we use throughout the paper.
Our wind farm model is based on
the notation and physical description 
from~\cite{Zhang2014}.  
Subsequently, we provide a literature review of 
existing mathematical
optimization methods for WFLO, focusing
on their similarities and differences
compared to our approach.

\subsection{The Wind Farm Model}
\begin{figure}[t]
	\centering
	\includegraphics[scale = 0.9]{field_model.pdf}

	\caption{Wind farm model: proximity constraints, wakes, wind direction.}\label{fig:field_model}
\end{figure}

\paragraph{Wind Farm as a Grid} We consider a square shaped wind farm that we model 
as an $n\times n$ grid
with $n$ being the number of cells on each axis. 
The cells have an area of $c^2$ (in squared meters), i.e.,
if $c=100$ meters and $n=20$, the farm area will be $2km\times 2km$.
In this work, we assume that the terrain is flat and that 
we have no constraints related to the terrain.
Furthermore, we do not account for the influence of noise
on the wind farm's surroundings. The wind farm literature treats the more complex
farm models that capture non-flat terrain~\cite{song2015lazy,kuo2016wind}, multi-typed turbines~\cite{feng2017design}, noise effects~\cite{Zhang2014,sorkhabi2016impact,yin2014multi}, and complex objectives (including turbine installation  and maintenance costs)~\cite{lackner2007analytical}.
In this work, we focus on the novel quadratic approach for WFLO, and hence
consider the most abstract wind farm model. Yet, our model can be extended,
in the future, to include 
additional wind farm settings. \todo{perhaps discuss in future work}

Returning to the grid model, we assume that the number of homogeneous turbines is known 
in advance; we denote the number of turbines by $m$. To place these $m$ turbines, 
we must consider two types of interference effects, namely \emph{proximity constraints} and \emph{wakes}.
In what follows, we explain the two effects using a toy example of placing a single turbine on an 
$8\times8$ grid (see Figure~\ref{fig:field_model}).

\paragraph{Proximity Constraints} Turbines must be placed at least \emph{five rotor diameters} apart. Depending on the cell size, 
	we shall formulate the corresponding constraint as part of the optimization model. For example, assuming that a turbine is placed
	at the center of the cell (cell size of $100$ meters) and given that the rotor diameter is $40$ meters, 
	we must ensure that two turbines will not placed in adjacent cells. In Figure~\ref{fig:field_model}, 
	the red area around the turbine represents the cells in which we cannot place an additional turbine.

\paragraph{Wakes} Turbines influence each other through interference effects that are referred to as \emph{wakes}~\cite{jensen1983note,shakoor2016wake}. 
	 Wakes cause reduction of the effective power produced by a turbine due to upstream turbines that change
	 its airflow dynamics. In Figure~\ref{fig:field_model}, the wind is assumed to be horizontal. Therefore, the wake effect due
	 to placing the turbine expands as shown in the diagram: it will influence only turbines that will be placed in the orange cells.
	 In other words, if a turbine is placed in the orange area, its effective
	 power is reduced due to the upstream turbine. Moreover, wakes can occur between multiple turbines, and hence
	 multiple wakes must be taken into account for each location when modeling the farm. 
	 The reduction in power due to the wake effect is determined by three parameters: (1) the distance between the upstream
	 turbine and the cell for which we want to compute the effect (denoted by $x$ in Figure~\ref{fig:field_model}), (2) the radius of the cone's opening 
	 (denoted by $r$ in Figure~\ref{fig:field_model}), and (3) 
	 the superposition of the wind regimes that correspond to the farm's location.
	 This reduction effect can be pre-computed for each pair of cells using known
	 turbine specification and terrain conditions together with probabilistic assumptions on wind regimes~\cite{Zhang2014}. Note that turbines can adjust their direction according to the current wind direction.

\paragraph{Effective Power Calculations in the Presence of Wakes} 
Let $\mathcal{D}$ be the set of wind regimes. Every regime $d$ is a combination of wind direction and free speed. 
The probability of wind regime 
$d$ is given by $p_d$ with $\sum_{d \in \mathcal{D}}^{} p_d = 1$. In the experimental section (Section~\ref{sec:eval}),
we provide one example of a probability distribution 
function over $36$ wind regimes (Figure~\ref{fig:prob_wind}). In that example, 
a regime $d=(10deg, 12km/h)$ corresponds to a north-eastern wind blowing at the velocity $12km/h$ and the
probability of this regime is given by $0.008$. Let $u_{id}$ be the wind velocity at turbine $i = 1,\ldots, m$ for wind regime $d\in\mathcal{D}$. Turbines can turn in alignment with the direction of the current wind regime.

Next, we write 
the expected power of a wind farm with $m$ turbines~(\cite{Zhang2014}): \begin{equation} E_{SS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \frac{1}{3} u_{id}^3p_d. \end{equation}
Given that we must consider wakes, the velocity $u_{id}$ is not equivalent to the free wind speed of current wind regime.
The actual velocity at turbine $i$, $u_{id}$, can be computed using the following equation~(\cite{Zhang2014}):
\begin{equation}
u_{id} = u_{id\infty} \Bigg[1 - \sqrt{\sum_{j\in\mathcal{U}_{id}} \bigg( 1-\frac{u_{ijd}}{u_{id\infty}} \bigg)^2}  \Bigg], \label{eq:ss}
\end{equation} with $u_{id\infty}$ being the free wind speed (second component of regime $d$), the set of turbines $\mathcal{U}_{id}$ that are upstream to $i$ 
under wind regime $d$, and $u_{ijd}$ being the wake 
reduced speed at $i$ due to upstream turbine $j$. 
The reduced speed, $u_{ijd}$, is computed using the distance $x$ between the two turbines ($i,j$)  and turbine and terrain specifications that yield the cone opening radius, $r$. Furthermore, the set of upstream turbines, $\mathcal{U}_{id}$ is computed  
by rotating the grid w.r.t. to the current wind regime $d \in \mathcal{D}$ and computing which turbine $j$ is located upstream to turbine $i$ given $d$. The expression in Eq. (\ref{eq:ss}) is referred to as the \emph{sum-of-squares} (SS) model for 
the total expected power in presence of wakes~\cite{Zhang2014}. This is also known to be the most accurate analytical expression account for multiple wakes.

An alternative 
form for the expected power in the presence of multiple wakes is the linear superposition (LS) model: \begin{equation} \label{eq:ls}
E_{LS} = \sum_{i=1}^m \sum_{d\in\mathcal{D}} \Bigg(\frac{1}{3}u_{id\infty}^3 -\sum_{j\in\mathcal{U}_{id}} \frac{1}{3}(u_{id\infty}^3 - u_{ijd}^3)\Bigg).
\end{equation} Note that $E_{LS}$ has the same parameters as $E_{SS}$; the difference is in the function that ties these parameters together. 
The LS model is known to be less accurate compared to SS, yet as we will show in Section it 
results in a quadratic objective function in our WFLO formulations, which is a desired property from an optimization standpoint.

@@

The allocation of turbines in a grid-like wind farm was first considered 
as an optimization problem in~\cite{MOSETTI1994105}. The proposed method in~\cite{MOSETTI1994105}
searches for improved solutions using a 
genetic approach~\cite{davis1991handbook},
which is an approximate technique 
that does not aim to provide an optimal solutions,
but rather attempts to find a quick solution to WFLO. 
In follow-up papers, 
the WFLO problem was solved using \emph{exact}  
optimization methodologies~\cite{turner2014new,Zhang2014}. 
These methods guarantee that the returned solution is optimal with respect
to the specified objective (i.e., maximize total expected power). 
Indeed, the exact methods were shown to yield higher energy values, yet they often suffer from 
high computational cost and long runtimes. Therefore, one must strike a balance between optimality and computational requirements of the solution.  
Below, we provide an overview of the literature of the two approaches (exact vs. approximate) and relate our 
work to both through quadratic optimization models\footnote{We shall review 
only literature that considers the same wind farm model as the one presented earlier in this section}.

\paragraph{Exact Optimization Methods} 
\todo{Rephrase and add citations to exact methods}
The WFLO problem can be solved to optimality 
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming).
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. For the wind farm model
presented in the beginning of this section,
the two dominant works that apply exact optimization 
are~\cite{Zhang2014}, and~\cite{turner2014new}. 
The former builds on~\cite{donovan2005wind} and solves a MILP model for the linear superposition (LS) objective
and use constrained programming to find a solution for the sum of squares (SS) objective. The latter, uses a quadratic and linear approximations to the SS objective. Since the SS objective is considered to be more accurate, solutions that come from LS-based optimization 
are plugged into the SS objective.

The main advantage of exact techniques
is that they provide guarantees on solution quality. However, their disadvantage is
that they may take a very long time to execute before finding an optimal solution. 
In fact, we shall show that the WFLO problem is computationally hard, a result
that we have not yet encountered in the WFLO literature. Therefore,
when multiple designs are considered and a fast solution is of essence,
one may turn to approximate WFLO solutions.

\paragraph{Approximate Methods} \todo{tighten} In contrast to exact methods, approximation algorithms often provide   
quick and sufficiently accurate solutions. Several works solved 
WFLO using evolutionary (genetic) approaches~\cite{MOSETTI1994105,gonzalez2010optimization,grady2005placement}. These methods use two principles (cross-over and mutation)
to change existing solutions and improve them over time. Evolutionary methods do not guarantee termination (i.e., they may run forever), and when they do terminate (in practical cases). Hence, one must often
introduce stopping criteria and thus compromise on the quality of the attained solution~\cite{davis1991handbook}.
Furthermore, there is no natural way of introducing constraints into genetic methods, which
leads to a feasibility check for every solution.

The first attempt to address these limitations 
was made in a local search approach~\cite{ozturk2004heuristic}. 
The solution still does not guarantee optimality, 
yet it circumvents the termination and feasibility checking issues of the evolutionary
solutions. The local search attempts to apply either move, remove or add actions given an incumbent 
feasible solution. When a turbine is moved, removed or added, the new solution is evaluated. The procedure terminates 
when a stopping criteria is reached.
The main limitation of~\cite{ozturk2004heuristic} is that it cannot perform an action 
that leads to an infeasible state. Therefore, it quickly converges to a sub-optimal solution far
from the optimal one. This was experimentally shown in~\cite{rivas2009solving}. To overcome this limitation, the authors of~\cite{rivas2009solving} use simulated annealing (SA), which is a neighborhood search method that allows visits to infeasible solutions. This creates a well-connected solution space that allows an escape for local optima, and was shown to be superior to~\cite{ozturk2004heuristic} and to the genetic method proposed in~\cite{grady2005placement}. A drawback of the SA approach in~\cite{rivas2009solving} is its ad-hoc nature. The various model components must be tailored and geared towards WFLO. This means that small adjustments to our WFLO setting
 (e.g., adding noise considerations), would require major changes to the SA solution. 
Additional methods and experimental comparisons between various approximate algorithms are available in~\cite{samorani2013wind}.

In our work, we also use an SA approach (similarly to~\cite{rivas2009solving}). However,
our methodology is generic since it is based on quadratic programming models.
These models can be easily adjusted to additional constraints and components of the objective function. In addition, formulating the problem using quadratic programming  enables us to solve WFLO using advanced software exact optimization solvers (e.g., Gurobi optimizer) and approximate hardware solutions (e.g., Fujitsu's digital annealer for solving quadratic problems using SA).

\paragraph{Quadratic Models for Exact and Approximate WFLO}
\todo{send a clear message here. Perhaps move to next section}.

A benchmark study compared state of the art exact approaches from~\cite{Zhang2014} to 
approximate methods and found that the two perform in a comparable fashion, without large differences between
the attained energy~\cite{yang2019simulated}. Exact methods tend to perform better
in smaller WFLO benchmarks, while simulated annealing methods had slightly dominating results
for larger benchmarks. This leads to the conclusion that the two types of methods should be considered.

In this work, we use a common optimization model to solve WFLO both exactly and approximately, thus enjoying the  best of both worlds. 
The common optimization model is a quadratic programming representation of WFLO, 
which attempts to minimize a quadratic energy function with or without considering additional constraints.
Specifically, we consider two variations of the quadratic model. 
The first one is referred to as quadratic constrained optimization problems (QCOP). 
It enables maximizing a quadratic (LS-based) energy expression, while 
ensuring that WFLO constraints are maintained (turbine proximity and total number of turbines is $m$). 
The second model is a quadratic unconstrained binary optimization (QUBO), 
which

On the one hand, exact solvers for quadratic programs (QCOP and QUBO) such as Gurobi~\cite{}, 
use cutting edge
operations research methods to enhance performance.
On the other hand, simulated annealing methods have been recently embedded onto 
designated processing units. These units use quantum-inspired technology to solve
complex problems quickly and with minimal degradation in terms of the resulting solution~\cite{}.
Specifically, in this work, we shall use Fujitsu's digital annealer (DA) to solve WFLO. The DA
also requires a quadratic representation of the WFLO problem.

Both of these methods use 
quadratic optimization models, which we shall present in the next section (Section~\ref{sec:QUBO4WFLO}).

@@

In this part, 
we formulate the WFLO problem
using quadratic programming. Specifically, we consider two
quadratic optimization problems that represent WFLO: a
quadratic constrained optimization (QCOP) problem and 
a quadratic unconstrained binary optimization (QUBO) problem. 
The former represents WFLO characteristics (proximity and number of turbines)
as hard constraints, while the latter considers these characteristics as soft constraints that are added to its objective function.

First, we present WFLO's inputs and decision variables.
Subsequently, we present the two quadratic formulations of WFLO
and discuss the usefulness of each formulation in the context
of advanced optimization software and hardware.

\subsection{WFLO Inputs and Decision Variables} We consider the wind farm model that we presented in Section~\ref{sec:related}) and follow the notation from~\cite{Zhang2014}. Let $x_i$ be a binary decision variable that represents whether a 
turbine is positioned at location 
$i \in \mathcal{N}$ with $\mathcal{N}$ being the set 
of $k = n^2$ possible turbine locations (cells in the grid). We 
write $x$ as shorthand for $x = (x_1,\ldots, x_k)$. Furthermore,
we denote by $\mathcal{E} \subseteq \mathcal{N}\times \mathcal{N}$
the set of location pairs that cannot simultaneously host turbines 
due to proximity constraints. The set $\mathcal{E}$ can be pre-computed 
based on problem specifications. For example,
it is common to assume that two turbines must be placed at least five rotor diameters apart.

Let $u_{id}$ and $u_{id\infty}$ 
be the wind speeds at turbine location $i \in \mathcal{N}$ for wind regime $d \in \mathcal{D}$
with and without interference from other turbines due to wake effects, respectively. 
Note that here we refer to $i$ as a potential turbine location and not the $i$th turbine as we did in Section~\ref{sec:related}.

Next, we denote by $\mathcal{U}_{id} \subseteq \mathcal{N}$ 
the set of downstream 
turbine locations for wind regime $d$
given that a turbine is placed at cell $i$ (i.e., turbines placed in these locations 
will be influenced by a turbine placed in $i$). Finally,
we denote by $u_{ijd}$ the wind speed at location $j$ due to a single wake from upstream turbine at location $i$ with $j \in \mathcal{U}_{id}$. 
The following function  
represents the total expected energy of a wind farm given placement decisions 
$x$: \begin{equation}
f(x) = \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j).   \label{ObjFunc}\\
%&s.t.:& \nonumber\\
%&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
%&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
%&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{equation} The objective function corresponds to the linear superposition (LS)
expression (Eq.~\ref{eq:ls}). We are now ready to provide the two quadratic programs to represent WFLO.

@@

To 
maximize $f(x)$ and satisfy the total number of turbines and proximity 
constraints we write
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}.
\end{eqnarray} The QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{}. 
So far, in~\cite{Zhang2014} and~\cite{donovan2005wind}, equivalent (linear) problems were formulated and solved using 
exact solvers. However, these solvers should only be used if the problem is computationally hard.
Prior to this work, we are unaware of a complexity result that analyzes LS-based WFLO. In what follows, we provide
a theorem that proves the computational complexity of the QCOP. 
\begin{mythm}
	The computational complexity of LS-based WFLO is $\mathcal{NP}$-hard. 
\end{mythm}
\begin{myproof}
We prove by showing that a special case of the WFLO problem (Eq.~(\ref{QCOP}) is
$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
are dropped and we get the HSP, which was proven to be $\mathcal{NP}$-hard in~\cite{billionnet2005different}.
	\end{myproof} To overcome this limitation associated with the computational hardness of WFLO,
we reformulate the QCOP into a QUBO problem. This allows us to use a specialized optimization hardware that uses
simulated annealing in order to solve QUBO problems quickly.

@@

\paragraph{WFLO Instances}
\begin{figure}[t]
	\centering
	\includegraphics[scale = 0.3]{prob_wind_pdf.pdf}

	\caption{Wind regimes: x-axis is the angle (times 10 degrees), y-axis is the probability for wind regime, color corresponds to free wind speed.}\label{fig:prob_wind}
\end{figure}

@@

\begin{table}[ht]
